
### Price

The mean and quartiles of the price are all less than $200, demonstrating that the majority of airbnb in New York City (at least 75%) have a non-expensive nightly price. The cheapest listing is $10 and the most expensive is $10,000. Because of the wide range and the skewed distribution of price values, we take the logarithm of the price to spread out the data

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}

# Basic histogram of price
ggplot(data=air.bnb.dataset,aes(air.bnb.dataset$price+1)) + 
  geom_histogram(boundary=0, bins = 10) + 
  scale_x_log10() +
  xlab("log(Nightly Price)")  + 
  ggtitle("Histogram of log of Nightly Price")

```

**Price Summary Statistics**
```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Summary statistics of price
summary(air.bnb.dataset$price)
```

Box plots of nightly price with respect to NYC borough demonstrate that the most expensive airbnb are in Manhattan, as one would expect.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Plot of NYC borough & price
ggplot(air.bnb.dataset, aes(x = neighbourhood_group_cleansed, y = log(price+1))) + 
  geom_boxplot()+ylab("Log of Price") + xlab("Borough")
```

There are 83 listings whose nightly price is \$30 or less. Likewise, there are 87 listings whose nightly price is at least \$1,500. The least expensive listings are mostly located in Brooklyn neighborhoods such as Bushwick, Bedford-Stuyvesant, and Crown Heights. Over 80% of the most expensive listings are located in Manhattan. 

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Plot locations of cheapest & most expensive locations
Listings.Cheap.Loc <- air.bnb.dataset[which(air.bnb.dataset$is_location_exact=="t" & air.bnb.dataset$price <= 30),c("latitude","longitude")]
No_Cheap <- nrow(Listings.Cheap.Loc)
Listings.Pricey.Loc <- air.bnb.dataset[which(air.bnb.dataset$is_location_exact=="t" & air.bnb.dataset$price >= 1500),c("latitude","longitude")]
No_Pricey <- nrow(Listings.Pricey.Loc)
Listings.Cheap.Loc <- cbind(Listings.Cheap.Loc,rep("Cheapest",No_Cheap))
names(Listings.Cheap.Loc)[3] <- "Cost"
Listings.Pricey.Loc <- cbind(Listings.Pricey.Loc,rep("Priciest",No_Pricey))
names(Listings.Pricey.Loc)[3] <- "Cost"
Listings.AllPrices <- rbind(Listings.Cheap.Loc,Listings.Pricey.Loc)

map <- get_map(location = 'New York City, NY', zoom = 11)
mapPoints <- ggmap(map) +
   geom_point(aes(x = longitude, y = latitude), data = Listings.AllPrices, alpha = .5) +
  facet_grid(. ~ Cost)

mapPoints
```



### Ratings 

Most airbnb, regardless of property type, tend to have very high reviews.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
layout(c(1,1))

ggplot(air.bnb.dataset, aes(x=property_type, y=review_scores_rating)) +
  geom_boxplot()+theme(axis.text.x=element_text(angle=90,size=12))+labs(x="Property Type",y="Review Score")

```

#### Ratings and Price

Price is not a strong indicator of the review rating, and the two variables appear to be independent.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Basic scatter plot of price & rating
ggplot(air.bnb.dataset, aes(x=log(price + 1), y=review_scores_rating)) +
  geom_point(shape=1)+labs(x="log(Nightly Price)",y="Review Score")

```

### Distance to Nearest Subway

Most listings are within 445 meters from a subway station. The listing closest to a subway station is within one meter, likely in a building right above it. The listing farthest from a subway station is 21,380 meters (~13 miles) away.

**Distance to Subway Summary Statistics**
```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Summary statistics & histogram of distance to nearest subway
summary(air.bnb.dataset$distance_in_meters)
```
```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
ggplot(data=air.bnb.dataset,aes(air.bnb.dataset$distance_in_meters+1)) + 
  geom_histogram(boundary=0, bins = 10) + 
  scale_x_log10() +
  xlab("log(Distance)")  + 
  ggtitle("Histogram of log of Distance to Nearest Subway Station")
```

There is no clear relationship between the distance of a listing to a subway station and that listing's price or review score.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
ggplot(air.bnb.dataset, aes(x=log(distance_in_meters + 1), y=log(price+1))) +
  geom_point(shape=1)+labs(x="log(Distance)",y="log(Nightly Price)")

# Basic scatter plot of distance to nearest subway & review score
ggplot(air.bnb.dataset, aes(x=log(distance_in_meters + 1), y=review_scores_rating)) +
  geom_point(shape=1)+labs(x="log(Distance)",y="Review Score")
```


### SuperHosts

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Basic summary statistics of superhost indicator variable
summary(air.bnb.dataset$host_is_superhost)
```

Only a small percentage of listings (~4%) are from "superhosts".

Brooklyn and Manhattan have the most superhosts, with both boroughs being home to nearly 90% of all superhosts in NYC.

Most superhosts in Brooklyn are in Williamsburg and Bedford-Stuyvesant. Within Manhattan, superhosts are more spread out, but tend to cluster in neighborhoods such as Harlem, Chelsea, Hell's Kitchen, and the East Village.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Table of NYC borough & superhosts
kable(table(airbnb$neighbourhood_group_cleansed,airbnb$host_is_superhost))
```



```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Plot locations of superhosts
Listings.Superhost.Loc <- air.bnb.dataset[which(air.bnb.dataset$is_location_exact=="t" & air.bnb.dataset$host_is_superhost == "t"),c("latitude","longitude")]


map <- get_map(location = 'New York City, NY', zoom = 10)
mapPoints <- ggmap(map) +
   geom_point(aes(x = longitude, y = latitude), data = Listings.Superhost.Loc, alpha = .5)+
  labs(title="Locations of Superhosts")
mapPoints
```


#### SuperHosts and Price

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Box plots of superhost & price (justification of Mann-Whitney Wilcoxon test)
ggplot(air.bnb.dataset, aes(x=host_is_superhost, y=log(price+1))) +
  geom_boxplot()+labs(x="Superhost Flag",y="log(Nightly Price)")
```

Box plots of the indicator variable of whether or not a host is a superhost against the log of the nightly price suggest that superhosts and non-superhosts charge similar prices for their listings. We can check this by running the Mann-Whitney Wilcoxon test.

#### SuperHosts and Subway Distance

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
# Box plots of superhost & distance to subway (justification of Mann-Whitney Wilcoxon test)
ggplot(air.bnb.dataset, aes(x=host_is_superhost, y=log(distance_in_meters+1))) +
  geom_boxplot()+labs(x="Superhost Flag",y="log(Distance)")
```

Box plots of the indicator variable of whether or not a host is a superhost against the log of the distance to the nearest subway station suggest that superhosts and non-superhosts have similar distributions for the distance to a subway stop. We can check this by running the Mann-Whitney Wilcoxon test.

### Correlations with Price

Numerical covariates tend to have a correlation of 0 with the price, suggesting the lack of a linear relationship between any single numerical variable and the price.

```{r, cache=TRUE, echo=FALSE, eval=TRUE, tidy=TRUE, warning=FALSE,message=FALSE}
NumericVariables <- c("distance_in_meters","Amenities.wirelessinternet","Couple"    ,"Host.Ver.email","Host.Ver.facebook","Host.Ver.googl","Host.Ver.jumio"             ,"Host.Ver.kba","Host.Ver.linkedin","Host.Ver.manualofflin","Host.Ver.manualonlin"   ,"Host.Ver.none","Host.Ver.phone","Host.Ver.photograph","Host.Ver.review"            ,"Host.Ver.sentid","Host.Ver.weibo","Amenities.aircondit","Amenities.breakfast"      ,"Amenities.buzzerwirelessintercom","Amenities.cabletv","Amenities.carbonmonoxidedetector","Amenities.cat","Amenities.dog","Amenities.doorman","Amenities.dryer"          ,"Amenities.elevatorinbuild","Amenities.essenti","Amenities.familykidfriend",        "Amenities.fireextinguish","Amenities.firstaidkit","Amenities.freeparkingonpremis"   
,"Amenities.gym","Amenities.heat","Amenities.hottub","Amenities.indoorfireplac"      ,"Amenities.internet","Amenities.kitchen","Amenities.otherpet" ,              "Amenities.petsallow","Amenities.petsliveonthisproperti","Amenities.pool"            ,"Amenities.safetycard","Amenities.shampoo","Amenities.smokedetector","Amenities.smokingallow","Amenities.suitableforev","Amenities.washer","Amenities.washerdry"    ,"Amenities.wheelchairaccess","calculated_host_airbnb_count","reviews_per_month", "number_of_reviews","review_scores_rating","review_scores_accuracy",           "review_scores_cleanliness","review_scores_checkin","review_scores_communication"    ,"review_scores_location","review_scores_value","minimum_nights"                ,"maximum_nights","availability_30","availability_365","guests_included", "bathrooms","bedrooms","beds","host_total_airbnb_count","host_response_rate","host_acceptance_rate")

airbnb.Numeric <- air.bnb.dataset[,(names(air.bnb.dataset) %in% NumericVariables)]
AllCor <- cor(airbnb.Numeric,airbnb.Numeric$price, use = "complete")

# Correlation between each numeric variable and price
kable(AllCor, digits=2)
```